---
title: "logistic_model"
author: "Jeremy Weidner"
date: "December 1, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notebook to begin fitting models on hotel data


JAEHEE
1) A schematic showing the relationship between the predictors and out come (include a list of variables and how they are measured)
- we will use EDA slides for this

2) A clear statement of the questions you are trying to answer.
- included in slides

3) Acsrtaining that histograms looks good, frequenicies within different cells look reasonable, if need be pool categories.
- also solved in EDA




JEREMY
4) A table of odds ratios with 95% CI and p-values that summarizes the results of your model.


5) Interpretation of odds ratios.



DAVID
6) Plot of interaction effect in the model and trying to interpret it. The fixed effect package should also give you plots for all predictors in the model.

7) mmp plots to see if model fits the data, as well as the pearson Chi-square test.

8) checking for outliers and influential points.

9) Some measure of pesudo R-square and accuracy of the model.




```{r cars}
# install.packages("effects")

library(effects)


hotels <- read.csv("hotels.csv")
head(hotels)
summary(hotels)
hotel_mod2 <- read.csv("hotel_mod2.csv")
 
```




```{r}
# 
# model_1 <- glm(is_canceled ~ lead_time + arrival_date_month + adults + children + babies + country + previous_cancellations + hotel + adr, family = binomial(), data = hotels)

# taking out country and children because they're not significant initially
model_2 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned, family = binomial(), data = hotels)
summary(model_2)
anova(model_2)

```

```{r}
hotels
```

```{r}
# Add meal 
model_3 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal, family = binomial(), data = hotels)
summary(model_3)
anova(model_2, model_3, test = "Chisq")
```

```{r}
# Add market_segment
model_4 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + market_segment, family = binomial(), data = hotels)
summary(model_4)
anova(model_3, model_4, test = "Chisq")
# Delete this one
```

```{r}
# Add deposit_type
model_5 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + market_segment + deposit_type, family = binomial(), data = hotels)
summary(model_5)
anova(model_4, model_5, test = "Chisq")
# Delete this one
```

```{r}
# Let's do some manipulations on days_in_waiting_list
dim(hotels)
If_wait <- rep(NA, 118984)
for (i in c(1:118984)) {
  if(hotels$days_in_waiting_list[i] == 0){
    If_wait[i] <- FALSE
  }
  else{
    If_wait[i] <- TRUE
  }
}
hotels$If_wait <- If_wait
# Add days_in_waiting_list
model_6 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + market_segment + deposit_type + If_wait, family = binomial(), data = hotels)
summary(model_6)
anova(model_5, model_6, test = "Chisq")
# To be decided......
```

```{r}
# Add customer_type
model_7 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + market_segment + deposit_type + If_wait + customer_type, family = binomial(), data = hotels)
summary(model_7)
anova(model_6, model_7, test = "Chisq")
```

```{r}
# required_car_parking_spaces
model_8 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type, family = binomial(), data = hotels)
summary(model_8)
anova(model_7, model_8, test = "Chisq")
# delete this one

```

```{r}
#After some testing our current model becomes:
curr_model <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type, family = binomial(), data = hotel_mod2)
```


```{r}
# Analysis on weather predictors
head(hotel_mod2)
hist(hotel_mod2$wdir)
hist(hotel_mod2$temp)
hist(hotel_mod2$maxt)
hist(hotel_mod2$visibility) # If add this?
hist(hotel_mod2$wspd)
hist(hotel_mod2$cloudcover)# If add this?
hist(hotel_mod2$mint)
hist(hotel_mod2$precip) # If add this?
hist(hotel_mod2$dew)
hist(hotel_mod2$humidity)
hist(hotel_mod2$precipcover) # If add this?
table(hotel_mod2$conditions)
```

```{r}
# Start to add weather predictors to our model
# Add wdir
curr_model1 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + wdir, family = binomial(), data = hotel_mod2)
summary(curr_model1)
anova(curr_model, curr_model1, test = "Chisq")
# Delete this one
```

```{r}
# Add temp
curr_model2 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp, family = binomial(), data = hotel_mod2)
summary(curr_model2)
anova(curr_model, curr_model2, test = "Chisq")
```

```{r}
# Add maxt
curr_model3 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + maxt, family = binomial(), data = hotel_mod2)
summary(curr_model3)
anova(curr_model2, curr_model3, test = "Chisq")
# Delete this one
```

```{r}
# Add visibility
curr_model4 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + visibility, family = binomial(), data = hotel_mod2)
summary(curr_model4)
anova(curr_model2, curr_model4, test = "Chisq")
# Some data is missed, delete this for now
```

```{r}
# Add wspd
curr_model5 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + wspd, family = binomial(), data = hotel_mod2)
summary(curr_model5)
anova(curr_model2, curr_model5, test = "Chisq")
# Delete this one
```

```{r}
# Add cloudcover
curr_model6 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + cloudcover, family = binomial(), data = hotel_mod2)
summary(curr_model6)
anova(curr_model2, curr_model6, test = "Chisq")
# Delete this one
```

```{r}
# Add mint
curr_model7 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + mint, family = binomial(), data = hotel_mod2)
summary(curr_model7)
anova(curr_model2, curr_model7, test = "Chisq")
```

```{r}
# Add precip
curr_model8 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + mint + precip, family = binomial(), data = hotel_mod2)
summary(curr_model8)
anova(curr_model7, curr_model8, test = "Chisq")
# Delete this one
```

```{r}
# Add dew
curr_model9 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + mint + dew, family = binomial(), data = hotel_mod2)
summary(curr_model9)
anova(curr_model7, curr_model9, test = "Chisq")
# Delete this one
```

```{r}
# Add humidity
curr_model10 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + mint + humidity, family = binomial(), data = hotel_mod2)
summary(curr_model10)
anova(curr_model7, curr_model10, test = "Chisq")
# Delete this one
```

```{r}
# Add precipcover
curr_model11 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + mint + precipcover, family = binomial(), data = hotel_mod2)
summary(curr_model11)
anova(curr_model7, curr_model11, test = "Chisq")
# Delete this one
```

```{r}
# Add conditions
curr_model12 <- glm(is_canceled ~ arrival_date_season*hotel + lead_time + adults +  babies +  international_booking + Stay_length + previous_cancellations + adr + reassigned + meal + If_wait + customer_type + temp + mint + conditions, family = binomial(), data = hotel_mod2)
summary(curr_model12)
anova(curr_model7, curr_model12, test = "Chisq")
# Delete this one

# Let's use curr_model7 for now
used_model <- curr_model7
```

# odds ratio

```{r}


round(exp(cbind(Estimate=coef(model_2), confint(model_2))),2)

# round(exp(cbind(Estimate=coef(model_8), confint(model_8))),2)

round(exp(cbind(Estimate=coef(used_model), confint(model_8))),2)
```

# plot interaction effect

```{r}
# plot(allEffects(model_2, ask = FALSE))

plot(allEffects(model_2), ask=FALSE, main = "Interaction Effect: ")


```







# mmp plot
```{r}

library(car)
mmp(model_2)
# mmp(model_8)
mmp(used_model)
```



# pearson chi square
```{r}


print(paste("Pearson’s X^2", round(sum(residuals(model_2, type="pearson")^2),3)))

print(paste("Pearson’s X^2", round(sum(residuals(model_8, type="pearson")^2),3)))

print(paste("Pearson’s X^2", round(sum(residuals(used_model, type="pearson")^2),3)))

```

# outlier test & influential points

```{r}

outlierTest(model_2)

residualPlots(model_2)

influencePlot(model_2)


```

































































































































































































































































